---
title: "Indv_Proj"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = T, message=FALSE, warning=FALSE)
```

<style type="text/css">

body{ /* Normal  */
      font-size: 12px;
  }
td {  /* Table  */
  font-size: 8px;
}
h1.title {
  font-size: 38px;
  color: DarkRed;
}
h1 { /* Header 1 */
  font-size: 28px;
  color: DarkBlue;
}
h2 { /* Header 2 */
    font-size: 22px;
  color: DarkBlue;
}

code.r{ /* Code block */
    font-size: 12px;
}
pre { /* Code block - determines code spacing between lines */
    font-size: 14px;
}
</style>

```{r}
library(dataQualityR) #for data cleaning
library(e1071) #SVM Training & Testing Models
library(mice) #deal with missing data
library(corrplot) #plot for correlation matrix
library(ggplot2) #visualization and plots
library(ggpubr) #customizing ggplot2
library(scales) #graphical scales map data to aesthetics
library(caret) #deal with Classification And REgression Training CART
library(dplyr) #data manipulation: filter and arrange
library(tidyverse) #data manipulation
library(sf) #plot mapping
library(gganimate) #static visualization
library(MASS) #deal with data set
library(VIM) #tools for the visualization of missing or imputed values
library(glmnet) #Lasso and Elastic-Net Regularization
```

# Outline:
1. Introduction - Business Context
2. Data Preparation
2.1. Data collection
2.2. Data Cleaning
2.3. Data transformation
2.4. Correlation Analysis
3. Exploratory Data Analysis - EDA
4. Modeling
5. Evaluation

# 1. Introduction

# 2. Data Preparation
- Fixed Acidity: are non-volatile acids that do not evaporate readily
- Volatile Acidity: are high acetic acid in wine which leads to an unpleasant vinegar taste
- Citric Acid: acts as a preservative to increase acidity. When in small quantities, adds freshness and flavor to wines
- Residual Sugar: is the amount of sugar remaining after fermentation stops. The key is to have a perfect balance between - sweetness and sourness. It is important to note that wines > 45g/ltrs are sweet
- Chlorides: the amount of salt in the wine
- Free Sulfur Dioxide: it prevents microbial growth and the oxidation of wine
- Total Sulfur Dioxide: is the amount of free + bound forms of SO2
- Density: sweeter wines have a higher density
- pH: describes the level of acidity on a scale of 0–14. Most wines are always between 3–4 on the pH scale
- Alcohol: available in small quantities in wines makes the drinkers sociable
- Sulphates: a wine additive that contributes to SO2 levels and acts as an antimicrobial and antioxidant
- Quality: which is the output variable/predictor

## 2.1. Data collection

## 2.2. Data cleaning

- Data reading
```{r}
df<-read.csv("winequality-red.csv")
dim(df)
```
- Data overview
```{r}
str(df)
summary(df)
```

- having a data report for cat and num 
```{r}
num.file <- paste(getwd(), "/dqames_num.csv", sep= "")
cat.file <- paste(getwd(), "/dqames_cat.csv", sep= "")
checkDataQuality(data= df, out.file.num= num.file, out.file.cat= cat.file)
```


- Checking missing values
```{r}
aggr_plot <- aggr(df, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(df), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))
```

## 2.4. Correlation analysis
- Build correlation and order by decreasing
```{r}
set.seed(123)
library(dplyr)
library(tidyr)

cor(df) %>%
  as.data.frame() %>%
  mutate(var1 = rownames(.)) %>%
  gather(var2, value, -var1) %>%
  arrange(desc(value)) %>%
  group_by(value) %>%
  filter(row_number()==1)
```

```{r}
dfcor <- cor(df)
corrplot(dfcor, method = "color", addCoef.col = "black",number.cex = .6,
         tl.col = "black", tl.srt = 90, diag = FALSE)
```

- Build correlation atts with Quality variable
```{r}
dfcor <- cor(df)
quality_cor <- dfcor[,12]
absoutcome_cor <- abs(quality_cor)
head(absoutcome_cor[order(absoutcome_cor, decreasing = TRUE)],12)
```

# 3. Exploratory Data Analysis - EDA

## 3.1. Overall Quality
```{r}
q1 <- ggplot(df, aes(quality))+ 
  geom_histogram() + 
  labs(title = "Histogram of quality") + 
  theme(plot.title=element_text(hjust=0.5)) +
  geom_vline(aes(xintercept=mean(quality)), color="blue", linetype="dashed", size=1) +
  geom_text(aes(x=5.6, label="Mean Value", y=400), colour="red", angle=90, vjust = 1.2, text=element_text(size=11))
q1

q2 <- ggplot(df, aes(sample=quality)) +
  stat_qq(color="dodgerblue4") + 
  stat_qq_line(color="red") +
  scale_y_continuous(labels=function(y){y/10^6}) +
  labs(title="QQ Plot for quality", y="Ordered Values") +
  theme(plot.title=element_text(hjust=0.5))
q2
```

## 3.2. Plotting the marginal distributions of key numerical quantities of interest

```{r}
p1 <- ggplot(df, aes(x=fixed.acidity)) + 
  geom_density()
p1 + geom_vline(aes(xintercept=mean(fixed.acidity)),
            color="blue", linetype="dashed", size=1)

p2 <- ggplot(df, aes(x=volatile.acidity)) + 
  geom_density()
p2 + geom_vline(aes(xintercept=mean(volatile.acidity)),
            color="blue", linetype="dashed", size=1)

p3 <- ggplot(df, aes(x=citric.acid)) + 
  geom_density()
p3 + geom_vline(aes(xintercept=mean(citric.acid)),
            color="blue", linetype="dashed", size=1)

p4 <- ggplot(df, aes(x=residual.sugar)) + 
  geom_density()
p4 + geom_vline(aes(xintercept=mean(residual.sugar)),
            color="blue", linetype="dashed", size=1)

p5 <- ggplot(df, aes(x=chlorides)) + 
  geom_density()
p5 + geom_vline(aes(xintercept=mean(chlorides)),
            color="blue", linetype="dashed", size=1)

p6 <- ggplot(df, aes(x=free.sulfur.dioxide)) + 
  geom_density()
p6 + geom_vline(aes(xintercept=mean(free.sulfur.dioxide)),
            color="blue", linetype="dashed", size=1)

p7 <- ggplot(df, aes(x=total.sulfur.dioxide)) + 
  geom_density()
p7 + geom_vline(aes(xintercept=mean(total.sulfur.dioxide)),
            color="blue", linetype="dashed", size=1)

p8 <- ggplot(df, aes(x=density)) + 
  geom_density()
p8 + geom_vline(aes(xintercept=mean(density)),
            color="blue", linetype="dashed", size=1)

p9 <- ggplot(df, aes(x=pH)) + 
  geom_density()
p9 + geom_vline(aes(xintercept=mean(pH)),
            color="blue", linetype="dashed", size=1)

p10 <- ggplot(df, aes(x=sulphates)) + 
  geom_density()
p10 + geom_vline(aes(xintercept=mean(sulphates)),
            color="blue", linetype="dashed", size=1)

p11 <- ggplot(df, aes(x=alcohol)) + 
  geom_density()
p11 + geom_vline(aes(xintercept=mean(alcohol)),
            color="blue", linetype="dashed", size=1)

ggarrange(p1, p2, p3, p4, nrow = 2, ncol =2)
ggarrange(p5, p6, p7, p8, nrow = 2, ncol =2)
ggarrange(p9, p10, p11, nrow = 2, ncol =2)
```

```{r}
b1 <- boxplot(df$fixed.acidity, col="slategray2", pch=19)
b2 <- boxplot(df$volatile.acidity, col="slategray2", pch=19)
b3 <- boxplot(df$citric.acid, col="slategray2", pch=19)
b4 <- boxplot(df$residual.sugar, col="slategray2", pch=19)
b5 <- boxplot(df$chlorides, col="slategray2", pch=19)
b6 <- boxplot(df$free.sulfur.dioxide, col="slategray2", pch=19)
b7 <- boxplot(df$total.sulfur.dioxide, col="slategray2", pch=19)
b8 <- boxplot(df$density, col="slategray2", pch=19)
b9 <- boxplot(df$pH, col="slategray2", pch=19)
b10 <- boxplot(df$sulphates, col="slategray2", pch=19)
b11 <- boxplot(df$alcohol, col="slategray2", pch=19)
```

## 3.3. Inspecting quality against numerical variables of interest

```{r}
g1 <- ggplot(df, aes(factor(quality), fixed.acidity, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "fixed.acidity", title = "Boxplot of Quality vs. fixed.acidity") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g1

g2 <- ggplot(df, aes(factor(quality), volatile.acidity, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "volatile.acidity", title = "Boxplot of Quality vs. volatile.acidity") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g2

g3 <- ggplot(df, aes(factor(quality), citric.acid, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "citric.acid", title = "Boxplot of Quality vs. citric.acid") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g3

g4 <- ggplot(df, aes(factor(quality), residual.sugar, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "residual.sugar", title = "Boxplot of Quality vs. residual.sugar") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g4

ggarrange(g1, g2, g3, g4, nrow = 2, ncol =2)

# It seems there’s positive relationship between citric acid and quality
# It seems there’s negative relationship between volatile acidity and quality
```

```{r}
g5 <- ggplot(df, aes(factor(quality), chlorides, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "Quality", y = "chlorides", title = "Boxplot of Quality vs. chlorides") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g5

g6 <- ggplot(df, aes(factor(quality), free.sulfur.dioxide, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "free.sulfur.dioxide", title = "Boxplot of quality vs. free.sulfur.dioxide") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g6

g7 <- ggplot(df, aes(factor(quality), total.sulfur.dioxide, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "total.sulfur.dioxide", title = "Boxplot of quality vs. total.sulfur.dioxide") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g7

g8 <- ggplot(df, aes(factor(quality), density, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "density", title = "Boxplot of quality vs. density") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g8

ggarrange(g5, g6, g7, g8, nrow = 2, ncol =2)

# It seems there’s negative relationship between density acid and quality
```

```{r}
g9 <- ggplot(df, aes(factor(quality), pH, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "pH", title = "Boxplot of Quality vs. pH") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g9

g10 <- ggplot(df, aes(factor(quality), sulphates, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "sulphates", title = "Boxplot of quality vs. sulphates") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g10

g11 <- ggplot(df, aes(factor(quality), alcohol, fill=factor(quality))) + 
  geom_boxplot() +
  labs(x = "quality", y = "alcohol", title = "Boxplot of quality vs. alcohol") + 
  theme(legend.position = 'none', plot.title = element_text(size = 9, hjust=0.5))
g11

ggarrange(g9, g10, g11, nrow = 2, ncol =2)

# It seems there’s positive relationship between alcohol and quality
# It seems there’s positive relationship between sulphates and quality
```

## 3.3. Inspecting relationship among numerical variables of interest
```{r}
s1 <- ggplot(df, aes(x=fixed.acidity, y=citric.acid)) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="fixed.acidity vs. citric.acid") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s1

s2 <- ggplot(df, aes(x=fixed.acidity, y=density)) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="fixed.acidity vs. density") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s2

s3 <- ggplot(df, aes(x=free.sulfur.dioxide, y=total.sulfur.dioxide)) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="free.sulfur.dioxide vs. total.sulfur.dioxide") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s3

s4 <- ggplot(df, aes(x=fixed.acidity, y=pH)) +
  geom_point(color="dodgerblue4",size=0.7) + 
  labs(title="fixed.acidity vs. pH") +
  geom_smooth(formula=y~x,method=lm, color="red") +
  theme(plot.title=element_text(hjust=0.5))
s4

ggarrange(s1, s2, s3, s4, nrow = 2, ncol =2)

# It seems there’s positive relationship between fixed.acidity and citric.acid
# It seems there’s positive relationship between fixed.acidity and density
# It seems there’s positive relationship between free.sulfur.dioxide and total.sulfur.dioxide
# It seems there’s negative relationship between fixed.acidity and pH
```

## 3.4. Inspecting 3D
```{r}
i1<- ggplot(df, aes(x=factor(round(alcohol)), y=citric.acid)) + 
  geom_boxplot(aes(colour = factor(quality))) +
  labs(title="Alcohol + Citric.Acid vs. Quality") + 
  theme(plot.title=element_text(hjust=0.5))
i1

i2 <- ggplot(df, aes(x=factor(round(alcohol)), y=volatile.acidity)) + 
  geom_boxplot(aes(colour = factor(quality))) +
  labs(title="Alcohol + Volatile.Acidity vs. Quality") + 
  theme(plot.title=element_text(hjust=0.5))
i2

i3 <- ggplot(df, aes(x=factor(round(alcohol)), y=chlorides)) + 
  geom_boxplot(aes(colour = factor(quality))) +
  labs(title="Alcohol + Chlorides vs. Quality") + ylim(0, 0.3)
  theme(plot.title=element_text(hjust=0.5))
i3
```

# 4. Modeling 
## 4.1. Modeling with top 5 variables

## 4.1. Model 1 with top 5 highest correlation with TotalIncidents
```{r}
lm0 <- lm(quality ~ alcohol + volatile.acidity +  sulphates + citric.acid + total.sulfur.dioxide, data = df)
summary(lm0)
```
```{r}
lm1 <- lm(quality ~ alcohol + volatile.acidity +  sulphates + total.sulfur.dioxide, data = df)
summary(lm1)
```

Using CV
```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10) #cv Cross-Validation
# Train the model
model1 <- train(quality ~ alcohol + volatile.acidity +  sulphates + total.sulfur.dioxide, data = df, method = "lm", trControl = train.control)
# Summarize the results
summary(model1)
print(model1)
```

## 4.2. LASSO
```{r}
library(glmnet)
x <- model.matrix(quality~., df)[,-1]
y <- df$quality
mod <- cv.glmnet(as.matrix(x), y, alpha=1)
```

To see the coefficients with the minimum cross-validation error
To see the coefficients with the "largest value of lambda such that error is within 1 standard error of the minimum:
```{r}
as.matrix(coef(mod, mod$lambda.min))
as.matrix(coef(mod, mod$lambda.1se))
```

You can also select any other value of lambda that you want. Coefficients that are 0 have been dropped out of the model
```{r}
CF <- as.matrix(coef(mod, mod$lambda.1se))
CF[CF!=0,]
```

Using Model 2 using above independent variables
```{r}
lm2 <- lm(quality ~ fixed.acidity + volatile.acidity + chlorides + total.sulfur.dioxide + sulphates + alcohol, data=df)
summary(lm2)
```

Using CV
```{r}
# Define training control
set.seed(123) 
train.control <- trainControl(method = "cv", number = 10) #cv Cross-Validation
# Train the model
model2<- train(quality ~ fixed.acidity + volatile.acidity + chlorides + total.sulfur.dioxide + sulphates + alcohol,data = df, method = "lm",
               trControl = train.control)
# Summarize the results
summary(model2)
print(model2)
```

## 4.3 Random Forest Model 3
```{r}
library(randomForest)
library(mlbench)
library(caret) # use createDataPartition() function 
set.seed(95014)

# partition 
#Create Evaluation Sets
set.seed(123)
n = nrow(df)
trainIndex = sample(1:n, size = round(0.7*n), replace=FALSE)
```

```{r}
#Creates training and test set from observations 
training = df[trainIndex,]
testing = df[-trainIndex,]

model3 <- randomForest(quality ~ ., training, mtry = 3, 
                         importance = TRUE, na.action = na.omit)
print(model3)

#Plot the error vs the number of trees graph 
plot(model3) 
```

```{r}
varImp(model3)
varImpPlot(model3,type=2)
```

```{r}
# obtain MSE as of last element in fit$mse
# which should match the output from printout
model3$mse[length(model3$mse)]
# take square root to calculate RMSE for the model
sqrt(model3$mse[length(model3$mse)])
```

```{r}
# now illustrate how to calculate RMSE on test data vs. training data
predValues <- predict(model3,testing)

# we can calculate it  directly 
sqrt(mean((testing$quality -predValues)^2)) #RMSE
mean(abs(testing$quality -predValues)) #MAE
```

Evaluation dataframe
```{r}
Model <- c("Model 1", "Model 2", "Model 3")
R_squared <- c(0.3479, 0.3546, 0.4850)
RMSE <- c(0.6549, 0.6515, 0.5843)
MAE <- c(0.5092899, 0.5063, 0.4222)
ml <- data.frame(Model, R_squared, RMSE, MAE)
```

Plot
```{r}
library(gridExtra)
p1 <- ggplot(ml, aes(Model, RMSE)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="RMSE") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
p2 <- ggplot(ml, aes(Model, R_squared)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="R-Squared") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
p3 <- ggplot(ml, aes(Model, MAE)) + geom_point(aes(colour = factor(Model), size = 4)) + labs(title="MAE") + theme(plot.title=element_text(hjust=0.5), axis.title.y = element_blank(),axis.title.x = element_blank(), legend.position="none")
grid.arrange(p2,p1,p3, ncol=3)
```

## 4.4. Clustering

Cleaning data
```{r}
df_new <- df
df_new$quality_group <- ifelse(df_new$quality < 5, "1", ifelse((df_new$quality >= 5) & (df_new$quality <= 6), "2", "3"))
df_new[,12]<-NULL # Order column looks like meaningless
df_new[,12]<-NULL # Order column looks like meaningless
str(df_new)
```

```{r}
df_new$quality_group <- as.numeric(df_new$quality_group)
str(df_new)
View(df_new)
```

```{r}
library(tidyverse)  # data manipulation
library(cluster)    # clustering algorithms
library(factoextra) # clustering algorithms & visualization
```

Standardize data
```{r}
scale(df_new)
head(df_new)
```

Elbow Method
```{r}
set.seed(123)
fviz_nbclust(df_new, kmeans, method = "wss")
```


Using K-means
```{r}
set.seed(123)
km <- kmeans(df_new, 2, nstart = 25)
df_new$cluster = km$cluster
View(df_new)
ggplot(df_new,aes(pH,alcohol,color=as.factor(cluster)))+geom_point()
ggplot(df_new,aes(pH,sulphates,color=as.factor(cluster)))+geom_point()
ggplot(df_new,aes(pH,total.sulfur.dioxide,color=as.factor(cluster)))+geom_point()
ggplot(df_new,aes(sulphates,total.sulfur.dioxide,color=as.factor(cluster)))+geom_point()
ggplot(df_new,aes(alcohol,sulphates,color=as.factor(cluster)))+geom_point()
ggplot(df_new,aes(alcohol,volatile.acidity,color=as.factor(cluster)))+geom_point()
```

